---
title: "A Corpus-Based Acoustic Analysis of Monophthong Vowels among Japanese Learners and Native Speakers of English - Part 2"
author: "Martin Schweinberger"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: word_document
---

# Introduction

This R Notebook shows how to access Praat from R and how to extract formant values from wav files and their corresponding Praat TextGrids. The tutorials is based on [Phonetics Tools in R](https://marissabarlaz.github.io/portfolio/phoneticstools/] by Marissa Barlaz).

## Preparation

install packages

```{r install, eval=F, message=F, warning=F}
# install
install.packages("tidyverse")
install.packages("here")
install.packages("adehabitatHR")
install.packages("lme4")
install.packages("sjPlot")
install.packages("report")
install.packages("flextable")
```

load packages

```{r load, message=F, warning=F}
library(tidyverse)
library(here)
library(adehabitatHR)
library(lme4)
library(sjPlot)
library(report)
library(flextable)
```


load data


```{r data, message=F, warning=F}
# load .rda data
fdat  <- base::readRDS(file = here::here("data", "fdat.rda"))
# inspect
nrow(fdat); head(fdat)
```




# Bhattacharyya's affinity 

function for extracting Bhattacharyya's affinity by type and target variety

```{r bafun, message=F, warning=F}
exba <- function(data, section, target, vwl1, vwl2){
  ba <- data %>%
  ungroup() %>%
  dplyr::filter(type == section,
                tvariety == target,
                label == vwl1 | label == vwl2)

  ba_formants <- ba %>%  dplyr::select(lobF1, lobF2)
  # extract vowels
  ba_vowel <- ba %>%  dplyr::select(vowel)
  # spatial data frame
  ba_spdf <- SpatialPointsDataFrame(ba_formants, ba_vowel)
  # calculate Bhattacharyya's affinity
  ba_ba <- kerneloverlap(ba_spdf, method = "BA")
  # result
  return(ba_ba[1,2])
}
```



## I vs i: 


```{r baius, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "us", vwl1 = "I", vwl2 = "i:")
# ENS
exba(data = fdat, section = "ENS", target = "us", vwl1 = "I", vwl2 = "i:")
```

```{r baigb, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "gb", vwl1 = "I", vwl2 = "i:")
# ENS
exba(data = fdat, section = "ENS", target = "gb", vwl1 = "I", vwl2 = "i:")
```

```{r Ii, message=F, warning=F}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = mean(lobF2),
                meanF1 = mean(lobF1)) %>%
  dplyr::filter(label == "i:" | label == "I") %>%
  ggplot(aes(x = lobF2, y = lobF1, color = vowel, fill = vowel)) +
  geom_point(alpha = 0.05) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  stat_ellipse(aes(group = label), level = .5) +
  geom_text(aes(x = meanF2, y = meanF1, 
                color = vowel, 
                label = vowel), size = 5) +
  #scale_color_viridis(discrete = TRUE) +
  scale_color_manual(values = c("red", "blue")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Formant 2 (Lobanov)", y = "Formant 1(Lobanov)")
ggsave(file = here("images", "hifront.png"), height = 5,  width = 5, dpi = 320)
```


## U vs u:

High-back 

```{r baius, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "us", vwl1 = "U", vwl2 = "u:")
# ENS
exba(data = fdat, section = "ENS", target = "us", vwl1 = "U", vwl2 = "u:")
```

```{r baigb, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "gb", vwl1 = "U", vwl2 = "u:")
# ENS
exba(data = fdat, section = "ENS", target = "gb", vwl1 = "U", vwl2 = "u:")
```


```{r, eval = F}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = mean(lobF2),
                meanF1 = mean(lobF1)) %>%
  dplyr::filter(label == "u:" | label == "U") %>%
  ggplot(aes(x = lobF2, y = lobF1, color = vowel, fill = vowel)) +
  geom_point(alpha = 0.05) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  stat_ellipse(aes(group = label), level = .95) +
  geom_text(aes(x = meanF2, y = meanF1, 
                color = vowel, 
                label = vowel), size = 5) +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right") +
  labs(x = "Hertz (mean Formant 2)", y = "Hertz (mean Formant 1)")
ggsave(file = here("images", "hiback.png"), height = 5,  width = 5, dpi = 320)
```


## Q vs O:

Mid-back 

```{r baius, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "us", vwl1 = "Q", vwl2 = "O:")
# ENS
exba(data = fdat, section = "ENS", target = "us", vwl1 = "Q", vwl2 = "O:")
```

```{r baigb, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "gb", vwl1 = "Q", vwl2 = "O:")
# ENS
#exba(data = fdat, section = "ENS", target = "gb", vwl1 = "Q", vwl2 = "O:")
```

```{r, eval = F}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = mean(lobF2),
                meanF1 = mean(lobF1)) %>%
  dplyr::filter(label == "O:" | label == "Q") %>%
  ggplot(aes(x = lobF2, y = lobF1, color = vowel, fill = vowel)) +
  geom_point(alpha = 0.1) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  stat_ellipse(aes(group = label), level = .95) +
  geom_text(aes(x = meanF2, y = meanF1, 
                color = vowel, 
                label = vowel), size = 5) +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right") +
  labs(x = "Hertz (mean Formant 2)", y = "Hertz (mean Formant 1)")
ggsave(file = here("images", "midback.png"), height = 5,  width = 5, dpi = 320)
```

## Ae vs e and E

Mid-front 

```{r baius, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "us", vwl1 = "E", vwl2 = "{")
# ENS
exba(data = fdat, section = "ENS", target = "us", vwl1 = "E", vwl2 = "{")
```

```{r baigb, message=F, warning=F}
# JPN
exba(data = fdat, section = "JPN", target = "gb", vwl1 = "e", vwl2 = "{")
# ENS
exba(data = fdat, section = "ENS", target = "gb", vwl1 = "e", vwl2 = "{")
```


```{r, eval = F}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = mean(lobF2),
                meanF1 = mean(lobF1)) %>%
  dplyr::filter(label == "E" | label == "e" | label == "{") %>%
  ggplot(aes(x = lobF2, y = lobF1, color = vowel, fill = vowel)) +
  geom_point(alpha = 0.05) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  stat_ellipse(aes(group = label), level = .95) +
  geom_text(aes(x = meanF2, y = meanF1, 
                color = vowel, 
                label = vowel), size = 5) +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right") +
  labs(x = "Hertz (mean Formant 2)", y = "Hertz (mean Formant 1)")
ggsave(file = here("images", "midfront.png"), height = 5,  width = 5, dpi = 320)
```


# Duration

```{r durtb, message=F, warning=F}
durtb <- fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::summarise(duration_mean = round(mean(duration), 3), 
                   duration_sd = round(sd(duration), 3),
                   freq = n())
# inspect
durtb
```


```{r}
durtb %>%
  dplyr::mutate(mean_sd = paste0(duration_mean, " (", duration_sd, ", N = ", freq, ")")) %>%
  dplyr::select(-duration_mean, -duration_sd, -freq) %>%
  tidyr::spread(type, mean_sd) %>%
  flextable::flextable()
```


```{r durplot, message=F, warning=F}
fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::summarise(duration = mean(duration)) %>%
  ggplot(aes(x = vowel, y = duration, fill = type)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  facet_grid( ~ tvariety) +
  theme_bw()
```



```{r}
diff_dat <- fdat %>%
  dplyr::group_by(type, tvariety, label) %>%
  dplyr::summarise(dur = mean(duration)) %>%
  dplyr::filter(label == "I" | 
                  label == "i:" | 
                  label == "U" | 
                  label == "u:" | 
                  label == "Q" | 
                  label == "O:" ) %>%
  tidyr::spread(label, dur) %>%
  dplyr::mutate(diff_i = `i:` - I,
                diff_u = `u:` - U,
                diff_o = `O:` - Q) %>%
  dplyr::select(-I, -U, -Q, -`i:`, -`u:`, -`O:`) %>%
  tidyr::gather(pair, diff, diff_i:diff_o) %>%
  dplyr::mutate(pair = dplyr::case_when(pair == "diff_i" ~ "Ii",
                                        pair == "diff_u" ~ "Uu",
                                        pair == "diff_o" ~ "Oo"))
# inspect
diff_dat
```


```{r}
diff_dat %>%
  ggplot(aes(x = pair, y = diff, fill = type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_grid(~tvariety)
```


# Overview tables


```{r tb1}
tb1 <- fdat %>%
  dplyr::group_by(type, tvariety) %>%
  dplyr::summarise(speakers = length(unique(fspeaker)),
                   vowel_tokens = n())
# inspect
tb1
```

```{r tb2}
tb2 <- fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::summarise(tokens = n(),
                   f1_raw = paste0(round(mean(F1), 1), 
                                   " (", round(sd(F1), 1), ")"), 
                   f1_lob = paste0(round(mean(lobF1), 3), 
                                   " (", round(sd(lobF1), 3),  ")"),
                   f1_bark = paste0(round(mean(barkF1), 2), 
                                    " (", round(sd(barkF1), 2),  ")"),
                   f2_raw = paste0(round(mean(F2), 1),
                                   " (", round(sd(F2), 1), ")"), 
                   f2_lob = paste0(round(mean(lobF2), 3),
                                   " (", round(sd(lobF2), 3), ")"), 
                   f2_bark = paste0(round(mean(barkF2), 2),
                                    " (", round(sd(barkF2), 2), ")"),
                   dur = paste0(round(mean(duration), 3), 
                                " (", round(sd(duration), 3), ")"))
# inspect
tb2
```



# Citation & Session Info

Schweinberger, Martin. `r format(Sys.time(), '%Y')`. A Corpus-Based Acoustic Analysis of Monophthongal Vowels among Japanese Learners and Native Speakers of English. Brisbane: The University of Queensland, School of Languages and Cultures. url: https://slcladal.github.io/praatrf.html (Version `r format(Sys.time(), '%Y.%m.%d')`).

@manual{schweinberger`r format(Sys.time(), '%Y')`praatrf,
  author = {Schweinberger, Martin},
  title = {A Corpus-Based Acoustic Analysis of Monophthongal Vowels among Japanese Learners and Native Speakers of English},
  note = {https://slcladal.github.io/praatrf.html},
  year = {`r format(Sys.time(), '%Y')`},
  organization = "The University of Queensland, School of Languages and Cultures},
  address = {Brisbane},
  edition = {`r format(Sys.time(), '%Y.%m.%d')`}
}


```{r}
sessionInfo()
```
