---
title: "A Corpus-Based Acoustic Analysis of Monophthong Vowels among Japanese Learners and Native Speakers of English - Part 2"
author: "Martin Schweinberger"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: word_document
---

# Introduction

This R Notebook shows how to access Praat from R and how to extract formant values from wav files and their corresponding Praat TextGrids. The tutorials is based on [Phonetics Tools in R](https://marissabarlaz.github.io/portfolio/phoneticstools/] by Marissa Barlaz).

## Preparation

install packages

```{r install,eval=F, message=F, warning=F}
# install
install.packages("tidyverse")
install.packages("here")
install.packages("viridis")
install.packages("phonR")
install.packages("vowels")
install.packages("ggridges")
install.packages("sm")
```

load packages

```{r load, message=F, warning=F}
library(tidyverse)
library(here)
library(viridis)
library(phonR)
library(vowels)
library(ggridges)
library(sm)
options("scipen" = 100, "digits" = 12) # suppress math annotation
```


load data


```{r data, message=F, warning=F}
# load .rda data
vdf  <- base::readRDS(file = here::here("data", "vdf.rda")) %>%
  dplyr::mutate(type = stringr::str_replace_all(file, "^[A-Z]{2,}_([A-Z]{3,3})_.*", "\\1"),
                speaker = stringr::str_replace_all(file, ".*_([0-9]{3,3})_.*", "\\1")) %>%
  dplyr::mutate(tgender = stringr::str_replace_all(path, ".*F[3-7]{1,1}_([fm]{1,1})_[a-z]{2,2}$", "\\1"),
                tformants = stringr::str_replace_all(path, ".*(F[3-7]{1,1})_[fm]{1,1}_[a-z]{2,2}$", "\\1"),
                tvariety = stringr::str_remove_all(path, ".*_"))
# inspect
nrow(vdf); head(vdf)
```


# Load Biodata

```{r bio, message=F, warning=F}
bio <- read.delim(here::here("data", "biodata_icnale.txt"), header = T, sep = "\t") %>%
  dplyr::filter(stringr::str_detect(Code, "JPN|ENS")) %>%
  dplyr::mutate(type = stringr::str_replace_all(Code, "^[A-Z]{2,}_([A-Z]{3,3})_.*", "\\1"),
                speaker = stringr::str_replace_all(Code, ".*_([0-9]{3,3}).*", "\\1"),
                Sex = tolower(Sex)) %>%
  dplyr::rename(gender = Sex)
bio[bio=="N/A"]<-NA
# inspect
head(bio)
```


# Join data sets

Join formant and biodata

```{r combine, message=F, warning=F}
fdat <- dplyr::left_join(vdf, bio, by = c("type", "speaker"))  %>%
  # filter out di- and triphthongs
  dplyr::mutate(nv = nchar(stringr::str_remove_all(label, ":"))) %>%
  dplyr::filter(nv == 1) %>%
  dplyr::filter(gender == tgender) %>%
  # remove  ref speakers with l1-var other than genam or gb
  dplyr::mutate(tvariety = dplyr::case_when(Country == "ENS_GBR" ~ "gb",
                                            Country == "ENS_CAN" ~ "us",
                                            Country == "ENS_USA" ~ "us",
                                             TRUE ~ tvariety)) %>%
  dplyr::filter(type == "JPN" | type == "ENS" & Country == "ENS_CAN" | Country == "ENS_GBR" | Country == "ENS_USA")
# inspect
head(fdat)
```


testing

```{r}
table(fdat$label, fdat$tvariety, fdat$type)
```




# Reference data

**GenAm**

[Yang, Byunggon. 1996. A comparative study of American English and Korean vowels produced by male and female speakers. *Journal of Phonetics* 24: 245–261.](http://fonetiks.info/bgyang/db/96jp.pdf)

```{r refame, message=F, warning=F}
ame <- read.delim(here::here("data/reference", "GenAm.txt"), sep = "\t", header = T) %>%
  dplyr::mutate(label = dplyr::case_when(item == "had" ~ "{",
                                         item == "heed" ~ "i:",
                                         item == "hid" ~ "I",
                                         item == "head" ~ "E",
                                         item == "hard" ~ "6",
                                         item == "whod" ~ "u:",
                                         item == "hood" ~ "U",
                                         item == "hoard" ~ "O:",
                                         item == "hud" ~ "V",
                                         item == "hod" ~ "Q",
                                         TRUE ~ "other"),
                gender = dplyr::case_when(sex == "female" ~ "f",
                                          TRUE ~ "m")) %>%
  dplyr::filter(label != "other") %>%
  dplyr::group_by(label, gender) %>%
  dplyr::summarise(target_f1 = mean(F1_Hz),
                   target_f2 = mean(F2_Hz),
                   target_f3 = mean(F3_Hz)) %>%
  dplyr::mutate(tvariety = "us")
# inspect data
head(ame)
```

**SSB**

[Deterding, David 2009. The Formants of Monophthong Vowels in Standard Southern British English Pronunciation. *Journal of the International Phonetic Association* 27, 1-2: 47-55. doi:10.1017/S0025100300005417](http://journals.cambridge.org/download.php?file=%2F7775_45AD5CD7849D3B7634A613DE6E61C294_journals__IPA_IPA27_1-2_S0025100300005417a.pdf&cover=Y&code=0639c34593b21cb18a9e7b504bdd5319)




```{r refbre, message=F, warning=F}
bre <- read.delim(here::here("data/reference", "SSB.txt"), sep = "\t", header = T) %>%
  dplyr::mutate(label = dplyr::case_when(item == "had" ~ "{",
                                         item == "heed" ~ "i:",
                                         item == "hid" ~ "I",
                                         item == "head" ~ "e",
                                         item == "hard" ~ "A:",
                                         item == "whod" ~ "u:",
                                         item == "hood" ~ "U",
                                         item == "hoard" ~ "O:",
                                         item == "hud" ~ "V",
                                         item == "hod" ~ "Q",
                                         TRUE ~ "other"),
                gender = dplyr::case_when(sex == "female" ~ "f",
                                          TRUE ~ "m")) %>%
  dplyr::filter(label != "other") %>%
  dplyr::group_by(label, gender) %>%
  dplyr::summarise(target_f1 = mean(F1_Hz),
                   target_f2 = mean(F2_Hz),
                   target_f3 = mean(F3_Hz)) %>%
  dplyr::mutate(tvariety = "gb")
# inspect data
head(bre)
```

Combine reference and obs data

```{r cref, message=F, warning=F}
ref <- rbind(ame, bre)
fdat <- dplyr::left_join(fdat, ref) %>%
  dplyr::filter(!is.na(target_f1)) %>%
  dplyr::mutate(edist = sqrt((F1-target_f1)^2 + (F2-target_f2)^2))
# inspect
head(fdat)
```


Determine best fit (f3:f7)


```{r fit, message=F, warning=F}
fdat <- fdat %>%
  dplyr::arrange(file, tmin, tvariety, tformants) %>%
  dplyr::group_by(file, tmin) %>% 
  dplyr::slice(which.min(edist))
# inspect
head(fdat)
```



remove words that only occur 1 time

```{r}
wrds <- names(table(fdat$word))[which(table(fdat$word) > 1)]
# reduce words
fdat <- fdat %>%
  dplyr::filter(word %in% wrds,
                nchar(word) < 9,
                word != "the") %>%
  # bark normalization
  dplyr::mutate(barkF1 = normBark(F1),
                barkF2 = normBark(F2)) %>%
  # lobanov normalization
  dplyr::ungroup() %>%
  dplyr::mutate(fspeaker = paste0(type, speaker, collaspe = "")) %>%
  dplyr::group_by(fspeaker) %>%
  dplyr::mutate(lobF1 = scale(F1),
                lobF2 = scale(F2)) %>%
  # remove superfluous columns
  dplyr::select(-path, -PTJ1..wds., -PTJ2..wds., -SMK1..wds., -SMK2..wds., -Self.Ev, 
                -Grade.Degree, -Acad..Genre, -VST, -CEFR, -INTM, -INSM, -INTM.INSM, 
                -INTM.INSM.1, -Primary, -Secondary, -College, -Inschool, -Outschool,
                -Reading, -Writing, -NS, -Presentation, -EssayW, -nv)
# inspect
head(fdat)
```



# (Semi-)Manual cross-evaluation

check for inconsistencies

```{r check, message=F, warning=F}
table(fdat$label, fdat$word)
```

exclude inconsistencies

```{r exclude, message=F, warning=F}
fdat <- fdat %>%
  dplyr::mutate(word = stringr::str_squish(word)) %>%
  dplyr::mutate(remove = dplyr::case_when(word == "--we"  ~ "remove",
                                          word == "ba"  ~ "remove",
                                          word == "bit" & label == "6" ~ "remove",
                                          word == "books" & label == "I" ~ "remove",
                                          word == "but" & label == "I" ~ "remove",
                                          word == "do" & label == "6" ~ "remove",
                                          word == "do" & label == "I" ~ "remove",
                                          word == "dust" & label == "V" ~ "remove",
                                          word == "get" & label == "I" ~ "remove",
                                          word == "gets" & label == "I" ~ "remove",
                                          word == "good" & label == "I" ~ "remove",
                                          word == "hea" ~ "remove",
                                          word == "peo" ~ "remove",
                                          word == "speak" & label == "I" ~ "remove",
                                          word == "sets" & label == "I" ~ "remove",
                                          word == "shops" & label == "I" ~ "remove",
                                          word == "so" & label == "U" ~ "remove",
                                          word == "that" & label == "I" ~ "remove",
                                          word == "that's" & label == "I" ~ "remove",
                                          word == "they're" ~ "remove",
                                          word == "they" ~ "remove",
                                          word == "that\u0081's" & label == "I" ~ "remove",
                                          word == "this" & label == "O:" ~ "remove",
                                          word == "to" ~ "remove",
                                          word == "too" & label == "I" ~ "remove",
                                          word == "too" & label == "6" ~ "remove",
                                          word == "two" & label != "u:" ~ "remove",
                                          word == "ways" & label == "I" ~ "remove",
                                          word == "we're" & label == "6" ~ "remove",
                                          word == "we're" & label == "6" ~ "remove",
                                          stringr::str_detect(word, "\\W") ~ "remove",
                                          TRUE ~ "keep")) %>%
  dplyr::filter(remove != "remove")
# inspect
table(fdat$label, fdat$word)
```



## Check durations

```{r}
fdat <- fdat %>%
  # duration
  dplyr::mutate(duration = tmax - tmin)
# inspect
head(fdat)
```

plot durations


```{r}
fdat %>%
  ggplot(aes(x = vowel, y = duration, fill = type)) +
  geom_boxplot()
```







```{r}
max_durs <- fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::summarise(du_min = min(duration),
                   du_max = max(duration),
                   du_ave = mean(duration),
                   du_med = median(duration),
                   du_sd = sd(duration),
                   du_2sd = sd(duration)*1.645) %>%
  dplyr::mutate(cutoff_duration = du_ave + du_2sd) %>%
  dplyr::select(type, tvariety, vowel, cutoff_duration)
# inspect
head(max_durs)
```


remove duration outliers

```{r}
fdat <- dplyr::left_join(fdat, max_durs) %>%
  dplyr::filter(duration <= cutoff_duration,
                ifelse(label != "A:", duration <= 0.4, duration <= 0.75))
# inspect
head(fdat); nrow(fdat)
```

```{r}
 fdat %>%
  ggplot(aes(y = vowel, x = duration, fill = type)) +
  geom_density_ridges(alpha = .5) +
  theme_ridges()
```


```{r}
dur_IiUu <-  fdat %>%
  dplyr::filter(label == "i:" | label == "I" | label == "u:" | label == "U") %>%
  ggplot(aes(y = vowel, x = duration, fill = type)) +
  geom_density_ridges(alpha = .5) +
  theme_ridges() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("orange", "gray"))
ggsave(here::here("images", "dur_IiUu.png"), height = 3,  width = 5, dpi = 320)
dur_IiUu
```


```{r}
fdat %>%
  ggplot(aes(x = vowel, y = duration, fill = type)) +
  geom_boxplot()
```



```{r}
fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::mutate(ldur = log(duration)) %>%
  ggplot(aes(x = vowel, y = ldur, fill = type)) +
  geom_boxplot() +
  facet_grid(~tvariety)
```


```{r}
fdat %>%
  dplyr::group_by(type, tvariety, vowel) %>%
  dplyr::mutate(ldur = log(duration)) %>%
  ggplot(aes(x = vowel, y = ldur, fill = type)) +
  geom_boxplot() +
  facet_grid(~tvariety)
```



# Remove outliers

```{r}
nrow(fdat)
fdat <- fdat %>% 
  dplyr::group_by(type, tvariety, label) %>%
  dplyr::mutate(densF1 = sm.density(barkF1,eval.points=barkF1)$estimate,
                densF2 = sm.density(barkF2,eval.points=barkF2)$estimate,
                qnt1F1 = quantile(densF1, .5),
                qnt1F2 = quantile(densF2, .5))
# inspect
nrow(fdat); head(fdat); summary(fdat$densF1)
```

```{r}
fdat <- fdat %>%
  # remove data point within the lowest 25 percent of densities among F1 and F2 values
  # but when label = A: and tvariety is gb
  dplyr::mutate(rmv = ifelse(label == "A:" & tvariety == "gb", "keep", "remove")) %>%
  dplyr::rowwise() %>%
  dplyr::filter(rmv == "keep" | densF1 > qnt1F1 | densF2 > qnt1F2) %>%
  # remove superfluous columns
  dplyr::select(-densF1, -densF2, -qnt1F1, -qnt1F2, -cutoff_duration, -rmv)
# inspect
head(fdat); nrow(fdat)
```






```{r sdrmv, eval = F}
fdat <- fdat %>%
  dplyr::filter(lobF1 > -1.96,
                lobF1 < 1.96,
                lobF2 > -1.96,
                lobF2 < 1.96)
```


# Remove GB target

```{r tvar}
fdat <- fdat %>%
  dplyr::filter(tvariety != "gb")
```

# Check density

check words

```{r wordplot, message=F, warning=F}
wordplot <- function(fdat, vwl){
  plt <- fdat %>%
    dplyr::filter(label == vwl) %>%
    dplyr::group_by(word) %>%
    dplyr::mutate(meanF2 = mean(lobF2),
                  meanF1 = mean(lobF1)) %>%
    # plot
    ggplot(aes(x = lobF2, y = lobF1, color = word, fill = word)) +
    geom_point(alpha = 0.05) +
    facet_grid( ~ type) +
    scale_x_reverse(position = "top") + 
    scale_y_reverse(position = "right") + 
    stat_ellipse(aes(group = label), level = .95) +
    geom_text(aes(x = meanF2, y = meanF1, 
                  color = word, 
                  label = word), size = 5) +
    scale_color_viridis(discrete = TRUE) +
    theme_minimal() +
    theme(legend.position = "right") +
    labs(x = "Formant 2 (Bark normalized)", y = "Formant 1 (Bark normalized)")
  return(plt)
  }
```


```{r wordplot, message=F, warning=F}
wordplot2 <- function(fdat, vwl){
  plt <- fdat %>%
    dplyr::filter(label == vwl) %>%
    dplyr::group_by(word) %>%
    dplyr::mutate(meanF2 = mean(lobF2),
                  meanF1 = mean(lobF1)) %>%
    # plot
    ggplot(aes(x = lobF2, y = lobF1)) +
    geom_density_2d_filled() +
    facet_grid( ~ type) +
    scale_x_reverse(position = "top") + 
    scale_y_reverse(position = "right") + 
    scale_fill_distiller(palette = "Blues", direction = 1) +
    geom_text(aes(x = meanF2, y = meanF1, 
                  label = word), size = 4, 
                  color = "gray90") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(x = "Formant 2 (Bark normalized)", y = "Formant 1 (Bark normalized)")
  return(plt)
  }
```


```{r}
wordplot3 <- function(fdat, vwl1, vwl2){
  plt <- fdat %>% 
    dplyr::filter(label == vwl1 | label == vwl2) %>%
    dplyr::group_by(word, label) %>%
    dplyr::mutate(meanF2 = mean(lobF2),
                  meanF1 = mean(lobF1)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(label) %>%
    dplyr::mutate(cF2 = mean(lobF2),
                  cF1 = mean(lobF1)) %>%
    # plot
    ggplot(aes(x = lobF2, y = lobF1)) +
  stat_density_2d(geom = "polygon",
                  aes(alpha = ..level.., fill = label), bins = 8)  +
    facet_grid( ~ type) +
    scale_x_reverse(position = "top") + 
    scale_y_reverse(position = "right") + 
    #scale_fill_distiller(palette = "Blues", direction = 1) +
    geom_text(aes(x = meanF2, y = meanF1, 
                  label = word, color = label), size = 4) +
    geom_text(aes(x = cF2, y = cF1, 
                  label = label), size = 6, color = "gray20") +
    theme_minimal() +
    theme(panel.grid.major = element_blank(), 
                  panel.grid.minor = element_blank(),
                  legend.position = "none") +
  scale_color_manual(values = c("orange3", "gray40")) +
  scale_fill_manual(values = c("orange", "gray")) +
    labs(x = "Formant 2 (Lobanov normalized)", y = "Formant 1 (Lobanov normalized)")
  return(plt)
  }
```


```{r}
pIi <- wordplot3(fdat = fdat, vwl1 = "I", vwl2 = "i:")
ggsave(here::here("images", "pIi.png"), height = 3,  width = 5, dpi = 320)
pIi
```

```{r}
pUu <- wordplot3(fdat = fdat, vwl1 = "U", vwl2 = "u:")
ggsave(here::here("images", "pUu.png"), height = 3,  width = 5, dpi = 320)
pUu
```





```{r word_I}
wordplot2(fdat = fdat, vwl = "I")
```

```{r word_i}
wordplot2(fdat = fdat, vwl = "i:")
```



```{r word_U}
wordplot2(fdat = fdat, vwl = "U")
```


```{r word_uu}
wordplot2(fdat = fdat, vwl = "u:")
```

```{r word_Q}
wordplot2(fdat = fdat, vwl = "Q")
```



```{r word_V}
wordplot2(fdat = fdat, vwl = "V")
```



```{r word_oo}
wordplot2(fdat = fdat, vwl = "O:")
```



```{r word_ae}
wordplot2(fdat = fdat, vwl = "{")
```



```{r word_E}
wordplot2(fdat = fdat, vwl = "E")
```



```{r word_e}
wordplot2(fdat = fdat, vwl = "e")
```

# Normalization

See [here](https://marissabarlaz.github.io/portfolio/vowelnormalization/)

Gender normalization

```{r}
# Using mean F3 of open vowels (vowels with F1>600) to approximate female F1&F2 to those of male
tempvowel <- fdat %>% 
  dplyr::filter(F1 > 600) %>% 
  dplyr::group_by(gender) %>%
  dplyr::mutate(meanF3 = mean(F3))
# extract k: k = ratio of Female F3 to Male F3
k <- max(tempvowel$meanF3) / min(tempvowel$meanF3)
# perform normalization
fdat <- fdat %>%
  dplyr::mutate(barkF1 = phonR::normBark(F1),
                barkF2 = phonR::normBark(F2)) %>%
  dplyr::mutate(normF1 = dplyr::if_else(gender == "f", F1*k, F1),
                normF2 = dplyr::if_else(gender == "f", F2*k, F2))
```

calculate distance from target vowel centroid for each token

```{r}
tcentroid <- fdat %>%
  dplyr::filter(type == "ENS") %>%
  dplyr::group_by(tvariety, vowel) %>%
  dplyr::summarise(cF1 = median(barkF1),
                   cF2 = median(barkF2))
# inspect
head(tcentroid)
```

combine

```{r}
fdat <- dplyr::left_join(fdat, tcentroid) %>%
  # euclidean distance
  dplyr::rowwise() %>%
  dplyr::mutate(ED = sqrt((barkF1-cF1)^2 + (barkF2-cF2)^2)) %>%
  # remove superfluous columns
  dplyr::select(-remove)
# inspect
head(fdat)
```

# Gramm. Status of Word

```{r}
gram <- c("be", "but", "did", "do", "get", "gets", "got", "had", "has", "he", "his", 
          "she", "that", "this", "too", "was", "we", "what", "who", "with")
fdat <- fdat %>%
  # gramm. category of word
  dplyr::mutate(WordType = ifelse(word %in% gram, "grammatical", "lexical"))
```




```{r, eval = T}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = median(barkF2),
                meanF1 = median(barkF1)) %>%
  ggplot(aes(x = barkF2, y = barkF1, color = vowel, fill = vowel, label = vowel)) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  #stat_ellipse(aes(group = label), level = .5) +
  #geom_point(alpha = .2) +
  geom_text(alpha = .2) +
  #geom_count() +
  #geom_text(aes(x = meanF2, y = meanF1, color = vowel, label = vowel), size = 5) +
  #scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right") +
  labs(x = "Formant 2 (Bark normalized)", y = "Formant 1 (Bark normalized))")
ggsave(file = here("images", "vowels_lob.png"), height = 5,  width = 5, dpi = 320)
```

```{r, eval = T}
fdat %>%
  dplyr::group_by(label) %>%
  dplyr::mutate(meanF2 = median(barkF2),
                meanF1 = median(barkF1)) %>%
  ggplot(aes(x = barkF2, y = barkF1)) +
  facet_grid(tvariety ~ type) +
  scale_x_reverse(position = "top") + 
  scale_y_reverse(position = "right") + 
  geom_density_2d_filled() +
  #geom_density2d() +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right") +
  labs(x = "Formant 2 (Lobanov)", y = "Formant 1 (Lobanov))")
ggsave(file = here("images", "vowels_dens.png"), height = 5,  width = 5, dpi = 320)
```


# Overview of the data

```{r tabluate, message=F, warning=F}
fdat %>%
  dplyr::group_by()
```



# Save data to disc

```{r savefdat, message=F, warning=F}
# save
base::saveRDS(fdat, file = here::here("data", "fdat.rda"))
# inspect
nrow(fdat); head(fdat)
```


# Citation & Session Info

Schweinberger, Martin. `r format(Sys.time(), '%Y')`. A Corpus-Based Acoustic Analysis of Monophthongal Vowels among Japanese Learners and Native Speakers of English. Brisbane: The University of Queensland, School of Languages and Cultures. url: https://slcladal.github.io/praatrf.html (Version `r format(Sys.time(), '%Y.%m.%d')`).

@manual{schweinberger`r format(Sys.time(), '%Y')`praatrf,
  author = {Schweinberger, Martin},
  title = {A Corpus-Based Acoustic Analysis of Monophthongal Vowels among Japanese Learners and Native Speakers of English},
  note = {https://slcladal.github.io/praatrf.html},
  year = {`r format(Sys.time(), '%Y')`},
  organization = "The University of Queensland, School of Languages and Cultures},
  address = {Brisbane},
  edition = {`r format(Sys.time(), '%Y.%m.%d')`}
}


```{r}
sessionInfo()
```
